Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis

### 2024-04-23 paper/JoPC/response_letter_2nd_round.tex

		Another criterion related to the control input we considered was the energy consumed to pre-heat the heating medium in the heater. Although the energy consumption may sound as a relevant criterion, this is not directly associated with the MPC's control inputs. The reason is that the consumed energy corresponds to the control involving the auxiliary proportional controller, which sets the electric power. Therefore, any energy minimization should be primarily addressed by fine tuning of this auxiliary controller. To conclude, we prefer not to include this criterion that is unrelated to the MPC controller and keep the analysis of the control inputs in the visual form of the profiles in Figure~10 in the manuscript.		
	}

	\comment{
		In addition, the authors only summarize the best case performance improvements in the conclusions. This can be a bit misleading to readers since the proposed method also performs worse (significantly) in some parts of the simulations. So summarizing average values would be more appropriate to give a better picture of the overall performance.
	}

	\answer{
		%We thank the Reviewer for pointing out the confusing communication of the control performance improvements. 
		Based on the results related to relative improvements reported in our answer to the Reviewer's comment No.2, we updated the section Conclusion which now reads as follows:

@@

		\begin{quote}
			To properly investigate the control results, the control performance was also judged quantitatively using a set of quality criteria. The self-tunable control approach outperformed the conventional control strategy handling just a single controller, i.e., non-tunable controller. \change{In average, the control performance criteria improved compared to the upper and lower boundary MPC respectively as follows: the squared-error-based criterion (SSE) reduced by 23\,\% and 83\,\%, the maximal overshoot/undershoot $\sigma_{\mathrm{max}}$ reduced by 64\,\% and 102\,\%, and the settling time $t_{\epsilon}$ reduced by 12\,\% and 244\,\%.}
		\end{quote}
	}

	\comment{
		Based on the responses to my previous comments, it appears that the applicability of the proposed self-tunable explicit MPC method is limited to "small-scale multivariable systems, with decoupled dynamics"? If so, please mention this clearly in the abstract.
	}

	\answer{
		We thank the Reviewer for pointing out the applicability of the proposed method. 
		If the controlled system is multivariable, not exclusively small-scale, and has a decoupled dynamics, it is generally suitable for the implementation of the self-tuning method.		
		We incorporated the Reviewer's suggestion into the Abstract as follows:

### 2024-06-03 paper/JoPC/main_final.tex
response last round

\newcommand{\change}[1]{\textcolor{black}{#1}}

@@

		\begin{abstract}
			%%max 250 words!			
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to \change{retune the controller subject to the changed working conditions.} Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable technique is suitable for SISO (single-input and single-output) systems and MIMO (multiple-input and multiple-output) systems with decoupled dynamics. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}

@@

	The heat exchangers in their numerous variants are integrated into many industrial plants as the heat transfer represents the crucial phenomena for all thermal energy applications~\cite{KV18}. \change{Simultaneously, the utility generation for heating or cooling is energy-demanding. From the control viewpoint, the controller design for the heat exchangers is a challenging task due to the necessity to take into account the nonlinear and asymmetric behavior of the device, i.e., different plant behavior when the temperature is increasing, in contrast to the behavior when the temperature is decreasing, see~\cite{RL20}.}

@@

	The applicability of model predictive control expanded with the parametric solution of the MPC optimization problem, known as explicit MPC~\cite{Bemporad_automatica}. As the MPC optimization problem is pre-solved offline, it does not need to be solved in the online phase, i.e., in real-time control. Instead, a piece-wise affine (PWA) control law is evaluated to apply the optimal control action in each control step.
	\change{The complexity of construction of the explicit MPC controller grows exponentially with the number of considered constraints. If the MPC design problem can be pre-solved explicitly offline, the consequent reduced online computational complexity makes the explicit MPC more suitable for practical industrial implementation.} Nevertheless, the explicit MPC is not tunable in default as the conventional approach in~\cite{Bemporad_automatica} considers the penalty matrices with fixed structure and values. \change{The inability to tune the explicit controller online can be a disadvantage due to varying operating conditions when the different setups of the controllers are beneficial.}

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical industrial-oriented implementation. \change{In numerous practical applications, the reference value of the controlled variable is changed and acquires values from a wide range of operating conditions. The use of different controller setups can help handle the plant's nonlinear behavior. The paper~\cite{self_tunable} presents a procedure of the self-tunable controller technique.} The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. \change{In the context of MPC, the aggressiveness is associated with the setup of the penalty matrices, as it determines the aggressiveness of the final control input. In general, higher penalization of the controlled states or control error in the cost function leads to more aggressive control actions. This process is analogous to increasing the proportional gain in the PID controller. On the contrary, higher penalization of the input variable leads to more sluggish control, e.g., see~\cite{Maciejowski_MPC}. In~\cite{self_tunable}, the MPC tuning} based on the distance from the steady-state operating point represented a way how to compensate for the system's nonlinear behavior.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, let us consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{I} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs, and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. These sets include the origin in their strict interiors. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}, \change{Q_\mathrm{y} \succeq 0}$ penalizes the squared control error, i.e., the deviation between the controlled output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}, \change{R \succ 0}$ penalizes the squared value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{I} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}, \change{Q_\mathrm{I} \succeq 0}$. \change{All the penalty matrices are considered to be diagonal due to the applicability of the self-tunable explicit MPC approach.}
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

@@

	The idea of approximated tunable explicit MPC comes from the work ~\cite{Klauco_tunable}, where the control action is calculated based on linear interpolation between two boundary control actions. These control actions result from evaluating two boundary explicit MPCs. The boundary explicit controllers \change{are constructed by solving the optimization problem having} the same structure and setup, except for one of the penalty matrices -- the tuned one. \change{Based on the specific control application, any penalty matrix can be chosen as the tuned parameter, i.e., this approach is applicable for any penalty matrix. The boundary penalty matrices follow the assumptions on the penalty matrices from Section 2.1 and are diagonal matrices such that $\lambda_{i,\mathrm{L}} \le \lambda_{i,\mathrm{U}}$, $\forall i = 1,\dots,s$, where $\lambda$ denotes the vector of eigenvalues of the penalty matrix, $s$ is the rank of the tuned penalty matrix, and $L$, $U$ denote the lower and upper boundary setup, respectively. }

	Let us consider the penalty matrices in the cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{I}(k) = (1-\rho(k)) \, Q_\mathrm{I,L} + \rho(k) \, Q_\mathrm{I,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance \change{by systematic tuning of all the penalty matrices. Systematic tuning involves selecting a specific penalty matrix and observing the control results by gradually increasing or decreasing the diagonal elements of the matrix. This process is then repeated for the remaining penalty matrices in a similar manner.}

@@

	Consider a single-input and single-output (SISO) system or a multiple-inputs and multiple-outputs (MIMO) system with completely decoupled pairs of the control inputs and the system outputs. 
	Then, the procedure of tuning the controller is based on evaluating the different operating points between the current value of the reference and the system steady-state value. This deviation is considered to scale the value of control action. First, the maximal admissible absolute value of the reference is defined. Analogous to the reference trajectory preview concept of MPC design, this value can be determined based on the general knowledge of the expected future reference values. Another suggestion is to set the maximal deviation $d_{\max}$ based on the constraints on system outputs: 	
	\begin{eqnarray}
		\label{eq:d_max}
		d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
	\end{eqnarray}	
	where \change{the symbol $\lvert.\rvert$, hereafter, denotes the element-wise absolute value}, $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable in the deviation form, i.e., zero (origin) corresponds to the system steady-state value. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation: 
	\change{ 
	\begin{eqnarray}
		\label{eq:rho_self_tunable}
		\rho (k) = \frac{\vert y_{\mathrm{ref}}(k) \vert}{d_{\max}}.
	\end{eqnarray}
	}
	Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a consequence, the parameter $\rho$ represents a way how to normalize the deviation from the steady-state value and is exploited to scale the control action or, implicitly, to tune the aggressiveness of the controller.

	\change{Note that the reference value must be reachable from the operating range to ensure that $0 \le \rho \le 1 $ holds. Otherwise, the interpolated control action would be the ``extrapolation'' leading to the loss of guarantees on the input or state constraints satisfaction, etc.}

@@

	Note that the relations in Eq.~\eqref{eq:rho_self_tunable} and Eq.~\eqref{eq:rho_max} operate with the absolute value of the reference. \change{It is not taken into account whether the reference value changed upwards or downwards with respect to the system steady-state value placed in the origin, i.e., whether the inequality $\Delta_\mathrm{ref}(k) = y_\mathrm{ref}(k)-y_\mathrm{ref}(k-1) > 0$ holds or $\Delta_\mathrm{ref}(k) < 0$. As many plants have nonlinear behavior with an asymmetric nature (different behavior when the process variable is rising or decreasing), the positivity or negativity of the reference change could be considered in the controller self-tuning procedure to improve the control performance.}

@@

	\change{In Eq.~\eqref{eq:rho_delta_ref}, it is suggested to increase the value of tuning parameter $\rho$ with increasing value of reference step change. Note, in this work, the larger value of the tuning parameter leads to adding more weight on the penalty matrices associated with the upper boundary controller, see Eq.~\eqref{eq:tunable_matrices}. If the opposite logic of controller tuning is requested, it is possible to adapt the tuning such that 
	\begin{eqnarray}
		\label{eq:tunable_R_2}
		&~& R(k) = \rho(k) \, R_\mathrm{L} + (1-\rho(k)) \, R_\mathrm{U}, \\
		\label{eq:tunable_Qx_2}
		&~& Q_\mathrm{I}(k) = \rho(k) \, Q_\mathrm{I,L} + (1-\rho(k)) \, Q_\mathrm{I,U}, \\
		\label{eq:tunable_Qy_2}
		&~& Q_\mathrm{y}(k) = \rho(k) \, Q_\mathrm{y,L} + (1-\rho(k)) \, Q_\mathrm{y,U},
	\end{eqnarray}
	hold. This change leads to adding more weight to the lower boundary controller with the increasing value of the tuning parameter $\rho$.}

@@

	The plant on which the control was implemented and analyzed is a laboratory-scaled counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. \change{The schematic of the plant is depicted in Figure~\ref{fig:HE_scheme}.} The heat exchanger is 90\,mm wide, 103\,mm long, and 160\,mm high. The heat exchange is performed between the \change{cold medium} (water) and the hot medium (water). \change{The cold medium as well as the heating medium are transported to the heat exchanger by two peristaltic pumps with flexible tubing from silicon rubber. The flow rate of the cold medium is constant, while the aim of control is to track the reference value of the outlet cold medium temperature. Therefore, the controlled variable is the \change{cold medium} temperature $T$ at the outlet of the heat exchanger. The inlet cold medium temperature was constant during the whole control, i.e., $T_\mathrm{C} = 19^{\circ}$\,C.} The temperature of the heated \change{cold medium} in the outlet stream was measured by the type K thermocouple. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium. The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. The maximal voltage 5\,V or 100\% corresponds to volumetric flow rate 11.5\,$\mathrm{ml}\,\mathrm{s}^{-1}$. For further technical specifications of the laboratory heat exchanger, the reader is referred to~\cite{pct23}. As heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy. The corresponding illustrative scheme of the implemented closed-loop control setup is in Figure~\ref{fig:CL}, where the ``Self-tuning'' block substitutes the more detailed scheme of the tuning procedure in Figure~\ref{fig:tuning}.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/HE}
		\caption[Heat exchanger Armfield Process Plant Trainer PCT23]{Laboratory heat exchanger Armfield Process Plant Trainer PCT23: \change{cold medium} pump\,(1), heating medium pump\,(2), \change{cold medium} tanks\,(3), heater for heating medium\,(4), heat exchanger\,(5).}
		\label{fig:HE}
	\end{center}
\end{figure}

>>>>

@@

		\begin{abstract}						
			The tunable approximated explicit model predictive control (MPC) comes with the benefits of real-time tunability without the necessity of solving the optimization problem online. This paper provides a novel self-tunable control policy that does not require any interventions of the control engineer during operation in order to retune the controller subject to the changed working conditions. Based on the current operating conditions, the autonomous tuning parameter scales the control input using linear interpolation between the boundary optimal control actions. The adjustment of the tuning parameter depends on the current reference value, which makes this strategy suitable for reference tracking problems. Furthermore, a novel technique for scaling the tuning parameter is proposed. This extension provides to exploit different ranges of the tuning parameter assigned to specified operating conditions. The self-tunable explicit MPC was implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The asymmetric behavior of the plant was compensated by tuning the controller's aggressiveness, as the negative or positive sign of reference change was considered in the tuning procedure. The designed self-tunable controller improved control performance by decreasing sum-of-squared control error, maximal overshoots/undershoots, and settling time compared to the conventional control strategy based on a single (non-tunable) controller. 
		\end{abstract}
		%
		%%%Graphical abstract
		%\begin{graphicalabstract}
			%%\includegraphics{grabs}
			%\end{graphicalabstract}
		%
		%%Research highlights
%		\begin{highlights}
%			 %3-5 highlights, every one should not exceed 85 characters including spaces
%			\item Self-tunable control technique does not need any intervention during control
%			\item Tuning based on operating conditions compensates for the nonlinear plant behavior
%			\item Implementation of the proposed tuning method on a laboratory heat exchanger 
%			\item Control performance improved compared to utilizing explicit MPC with constant setup
%		\end{highlights}

@@

	The heat exchangers in their numerous variants are integrated into many industrial plants as the heat transfer represents the crucial phenomena for all thermal energy applications~\cite{KV18}. Simultaneously, the utility generation for heating or cooling is energy-demanding. From the control viewpoint, the controller design for the heat exchangers is a challenging task due to the necessity to take into account the nonlinear and asymmetric behavior of the device, i.e., different plant behavior when the temperature is increasing, in contrast to the behavior when the temperature is decreasing, see~\cite{RL20}.

@@

	The applicability of model predictive control expanded with the parametric solution of the MPC optimization problem, known as explicit MPC~\cite{Bemporad_automatica}. As the MPC optimization problem is pre-solved offline, it does not need to be solved in the online phase, i.e., in real-time control. Instead, a piece-wise affine (PWA) control law is evaluated to apply the optimal control action in each control step.
	The complexity of construction of the explicit MPC controller grows exponentially with the number of considered constraints. If the MPC design problem can be pre-solved explicitly offline, the consequent reduced online computational complexity makes the explicit MPC more suitable for practical industrial implementation. Nevertheless, the explicit MPC is not tunable in default as the conventional approach in~\cite{Bemporad_automatica} considers the penalty matrices with fixed structure and values. The inability to tune the explicit controller online can be a disadvantage due to varying operating conditions when the different setups of the controllers are beneficial.

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical industrial-oriented implementation. In numerous practical applications, the reference value of the controlled variable is changed and acquires values from a wide range of operating conditions. The use of different controller setups can help handle the plant's nonlinear behavior. The paper~\cite{self_tunable} presents a procedure of the self-tunable controller technique. The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. In the context of MPC, the aggressiveness is associated with the setup of the penalty matrices, as it determines the aggressiveness of the final control input. In general, higher penalization of the controlled states or control error in the cost function leads to more aggressive control actions. This process is analogous to increasing the proportional gain in the PID controller. On the contrary, higher penalization of the input variable leads to more sluggish control, e.g., see~\cite{Maciejowski_MPC}. In~\cite{self_tunable}, the MPC tuning based on the distance from the steady-state operating point represented a way how to compensate for the system's nonlinear behavior.

@@

	Explicit model predictive control~\cite{Bemporad_automatica} utilizes a parametric solution of the model predictive control introducing its application range towards the systems with fast dynamics. Moreover, the explicit solution enables providing rigorous analysis and certification of the closed-loop system stability, constraints satisfaction, etc. As the explicit solution is available, real-time solving of the optimization problem in every control step is omitted. As this work deals with industrial-oriented implementation, let us consider the optimization problem in the following form:
	\begin{subequations}
		\label{eq:mpc_problem}
		\begin{eqnarray}
			\label{eq:mpc_problem_cost}
			\min_{u_0,u_{1},\ldots,u_{N-1}} &~& \! \sum_{k=0}^{N-1} \! \left( (y_k-y_\mathrm{ref})^{\intercal} Q_\mathrm{y} (y_k-y_\mathrm{ref}) + u_{k}^{\intercal} R u_{k} + x_{\mathrm{I},k}^{\intercal} Q_\mathrm{I} x_{\mathrm{I},k} \right)  \\
			\label{eq:mpc_problem_prediction_model_x}
			\mathrm{s.t.\!:} &~& \widetilde{x}_{k+1} = \widetilde{A}\,\widetilde{x}_{k} + \widetilde{B}\,u_{k}, \\
			\label{eq:mpc_problem_prediction_model_y}
			&~& y_{k} = \widetilde{C}\,\widetilde{x}_{k}, \\
			\label{eq:mpc_problem_input_constraints}
			&~& u_{k} \in \mathcal{U}, \\
			\label{eq:mpc_problem_state_constraints}
			&~& y_{k} \in \mathcal{Y}, \\
			\label{eq:mpc_problem_initial_coindition}
			&~& \widetilde{x}_{0} = \theta, \\
			\label{eq:mpc_problem_k_range}
			&~& k = 0,1,\ldots, N-1,
		\end{eqnarray}
	\end{subequations}
	where $k$ denotes the step of the prediction horizon $N$. 
	%
	To obtain the offset-free control results, the built-in integrator was included in the state-space model, e.g., see~\cite{Ruscio_MPC_integral}. 
	The prediction model in Eq.~\eqref{eq:mpc_problem_prediction_model_x}--\eqref{eq:mpc_problem_prediction_model_y} has the form of augmented linear time-invariant (LTI) system for a given augmented state matrix $\widetilde{A} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\widetilde{\mathrm{x}}}}$, augmented input matrix $\widetilde{B} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}} \times n_{\mathrm{u}}}$ and augmented output matrix $\widetilde{C} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\widetilde{\mathrm{x}}}}$. 
	%
	Variables $\widetilde{x} \in \mathbb{R}^{n_{\widetilde{\mathrm{x}}}}$, $u \in \mathbb{R}^{n_{\mathrm{u}}}$, $y \in \mathbb{R}^{n_{\mathrm{y}}}$ are vectors of corresponding augmented system states, control inputs, and system outputs, respectively. 
	%
	The sets $\mathcal{U} \subseteq \mathbb{R}^{n_{\mathrm{u}}}$, $\mathcal{Y} \subseteq \mathbb{R}^{n_{\mathrm{y}}}$ are convex polytopic sets of physical constraints on inputs and outputs, respectively. These sets include the origin in their strict interiors. The penalty matrix $Q_\mathrm{y} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}, Q_\mathrm{y} \succeq 0$ penalizes the squared control error, i.e., the deviation between the controlled output and output reference value $y_\mathrm{ref}$. The matrix $R \in \mathbb{R}^{n_{\mathrm{u}} \times n_{\mathrm{u}}}, R \succ 0$ penalizes the squared value of control inputs. 
	%
	The value of integrator is also penalized in the cost function with the penalty matrix $Q_\mathrm{I} \in \mathbb{R}^{n_{\mathrm{y}} \times n_{\mathrm{y}}}, Q_\mathrm{I} \succeq 0$. All the penalty matrices are considered to be diagonal due to the applicability of the self-tunable explicit MPC approach.
	The parameter $\theta \in \Theta$ in Eq.~\eqref{eq:mpc_problem_initial_coindition} represents the initial condition of the optimization problem for which it is parametrically pre-computed.

@@

	The idea of approximated tunable explicit MPC comes from the work ~\cite{Klauco_tunable}, where the control action is calculated based on linear interpolation between two boundary control actions. These control actions result from evaluating two boundary explicit MPCs. The boundary explicit controllers are constructed by solving the optimization problem having the same structure and setup, except for one of the penalty matrices -- the tuned one. Based on the specific control application, any penalty matrix can be chosen as the tuned parameter, i.e., this approach is applicable for any penalty matrix. The boundary penalty matrices follow the assumptions on the penalty matrices from Section 2.1 and are diagonal matrices such that $\lambda_{i,\mathrm{L}} \le \lambda_{i,\mathrm{U}}$, $\forall i = 1,\dots,s$, where $\lambda$ denotes the vector of eigenvalues of the penalty matrix, $s$ is the rank of the tuned penalty matrix, and $L$, $U$ denote the lower and upper boundary setup, respectively.

	Let us consider the penalty matrices in the cost function in Eq.~\eqref{eq:mpc_problem_cost}. The penalty matrices are scaled in the following way:
	\begin{subequations}
		\label{eq:tunable_matrices}
		\begin{eqnarray}
			\label{eq:tunable_R}
			&~& R(k) = (1-\rho(k)) \, R_\mathrm{L} + \rho(k) \, R_\mathrm{U}, \\
			\label{eq:tunable_Qx}
			&~& Q_\mathrm{I}(k) = (1-\rho(k)) \, Q_\mathrm{I,L} + \rho(k) \, Q_\mathrm{I,U}, \\
			\label{eq:tunable_Qy}
			&~& Q_\mathrm{y}(k) = (1-\rho(k)) \, Q_\mathrm{y,L} + \rho(k) \, Q_\mathrm{y,U},
		\end{eqnarray}
	\end{subequations}
	where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in Eq.~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in Eq.~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance by systematic tuning of all the penalty matrices. Systematic tuning involves selecting a specific penalty matrix and observing the control results by gradually increasing or decreasing the diagonal elements of the matrix. This process is then repeated for the remaining penalty matrices in a similar manner.

@@

	Consider a single-input and single-output (SISO) system or a system with completely decoupled pairs of the control inputs and the system outputs. 
	Then, the procedure of tuning the controller is based on evaluating the different operating points between the current value of the reference and the system steady-state value. This deviation is considered to scale the value of control action. First, the maximal admissible absolute value of the reference is defined. Analogous to the reference trajectory preview concept of MPC design, this value can be determined based on the general knowledge of the expected future reference values. Another suggestion is to set the maximal deviation $d_{\max}$ based on the constraints on system outputs: 	
	\begin{eqnarray}
		\label{eq:d_max}
		d_{\max} = \max(\vert y_{\min} \vert, y_{\max}),
	\end{eqnarray}	
	where the symbol $\lvert.\rvert$, hereafter, denotes the element-wise absolute value, $y_{\min}$ and $y_{\max}$ are respectively lower and upper bound on the output variable in the deviation form, i.e., zero (origin) corresponds to the system steady-state value. Using the information about the maximal possible deviation $d_{\max}$, the tuning parameter $\rho$ can be calculated as the ratio between the current reference value and the maximal deviation: 
	\begin{eqnarray}
		\label{eq:rho_self_tunable}
		\rho (k) = \frac{\vert y_{\mathrm{ref}}(k) \vert}{d_{\max}}.
	\end{eqnarray}
	Based on Eq.~\eqref{eq:rho_self_tunable}, the property $0 \le \rho \le 1$ holds, as $\vert y_{\mathrm{ref}} \vert \le d_{\max}$. As a consequence, the parameter $\rho$ represents a way how to normalize the deviation from the steady-state value and is exploited to scale the control action or, implicitly, to tune the aggressiveness of the controller.

	Note that the reference value must be reachable from the operating range to ensure that $0 \le \rho \le 1 $ holds. Otherwise, the interpolated control action would be the ``extrapolation'' leading to the loss of guarantees on the input or state constraints satisfaction, etc.

@@

	Note that the relations in Eq.~\eqref{eq:rho_self_tunable} and Eq.~\eqref{eq:rho_max} operate with the absolute value of the reference. It is not taken into account whether the reference value changed upwards or downwards with respect to the system steady-state value placed in the origin, i.e., whether the inequality $\Delta_\mathrm{ref}(k) = y_\mathrm{ref}(k)-y_\mathrm{ref}(k-1) > 0$ holds or $\Delta_\mathrm{ref}(k) < 0$. As many plants have nonlinear behavior with an asymmetric nature (different behavior when the process variable is rising or decreasing), the positivity or negativity of the reference change could be considered in the controller self-tuning procedure to improve the control performance.

@@

	In Eq.~\eqref{eq:rho_delta_ref}, it is suggested to increase the value of tuning parameter $\rho$ with increasing value of reference step change. Note, in this work, the larger value of the tuning parameter leads to adding more weight on the penalty matrices associated with the upper boundary controller, see Eq.~\eqref{eq:tunable_matrices}. If the opposite logic of controller tuning is requested, it is possible to adapt the tuning such that 
	\begin{eqnarray}
		\label{eq:tunable_R_2}
		&~& R(k) = \rho(k) \, R_\mathrm{L} + (1-\rho(k)) \, R_\mathrm{U}, \\
		\label{eq:tunable_Qx_2}
		&~& Q_\mathrm{I}(k) = \rho(k) \, Q_\mathrm{I,L} + (1-\rho(k)) \, Q_\mathrm{I,U}, \\
		\label{eq:tunable_Qy_2}
		&~& Q_\mathrm{y}(k) = \rho(k) \, Q_\mathrm{y,L} + (1-\rho(k)) \, Q_\mathrm{y,U},
	\end{eqnarray}
	hold. This change leads to adding more weight to the lower boundary controller with the increasing value of the tuning parameter $\rho$.

@@

	The plant on which the control was implemented and analyzed is a laboratory-scaled counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The schematic of the plant is depicted in Figure~\ref{fig:HE_scheme}. The heat exchanger is 90\,mm wide, 103\,mm long, and 160\,mm high. The heat exchange is performed between the cold medium (water) and the hot medium (water). The cold medium as well as the heating medium are transported to the heat exchanger by two peristaltic pumps with flexible tubing from silicon rubber. The flow rate of the cold medium is constant, while the aim of control is to track the reference value of the outlet cold medium temperature. Therefore, the controlled variable is the cold medium temperature $T$ at the outlet of the heat exchanger. The inlet cold medium temperature was constant during the whole control, i.e., $T_\mathrm{C} = 19^{\circ}$\,C. The temperature of the heated cold medium in the outlet stream was measured by the type K thermocouple. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium. The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. The maximal voltage 5\,V or 100\% corresponds to volumetric flow rate 11.5\,$\mathrm{ml}\,\mathrm{s}^{-1}$. For further technical specifications of the laboratory heat exchanger, the reader is referred to~\cite{pct23}. As heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy. The corresponding illustrative scheme of the implemented closed-loop control setup is in Figure~\ref{fig:CL}, where the ``Self-tuning'' block substitutes the more detailed scheme of the tuning procedure in Figure~\ref{fig:tuning}.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/HE}
		\caption[Heat exchanger Armfield Process Plant Trainer PCT23]{Laboratory heat exchanger Armfield Process Plant Trainer PCT23: cold medium pump\,(1), heating medium pump\,(2), cold medium tanks\,(3), heater for heating medium\,(4), heat exchanger\,(5).}
		\label{fig:HE}
	\end{center}
\end{figure}
